---
title: "lab6_mymarkdown"
author: "Sandra Fogg"
date: "2/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning  = FALSE)
```

Load Required Packages
```{r}
library(tidyverse)
library(here)
library(janitor)
library(raster)
library(sf)
library(tmap)
library(tmaptools)
library(gstat) # variogram
```

```{r}
# Look at it with base plot()
plot(gc_dem)

# Check CRS & bounds:
gc_dem@crs # Shows CRS: WGS84
gc_dem@extent # Shows extent (bounds)...notice that these seem odd (not units)

wgs84 = "+proj=longlat +datum=WGS84 +ellps=WGS84 +no_defs" # Just have this ready to copy/paste

# Reproject
gc_reproj = projectRaster(gc_dem, crs = wgs84, method = "bilinear")

# Then check: aha, now degrees we're used to
gc_reproj@extent

```

#### Crop the raster to a smaller area: 
```{r}
bounds <- as(extent(-112.4, -112.0, 36.1, 36.3), 'SpatialPolygons')

# Make the projection for "bounds" the same as for "gc_reproj":
crs(bounds) <- crs(gc_reproj)

# Then crop gc_reproj by the new bounds polygon: 
gc_crop <- crop(gc_reproj, bounds)

# Look at it:
plot(gc_crop)


```


Want to resample? Use `raster::aggregate()` to create lower res (larger cell) rasters.

See ?aggregrate (default is mean, `fact` is number of cells in each direction; can set 2 if wanting x/y to differ for aggregation)
```{r}
# Aggregate: 
gc_agg <- aggregate(gc_crop, fact = 10)

# Then look at it: 
plot(gc_agg)
```


Convert to GG
```{r}
gc_df <- as.data.frame(gc_agg, xy = TRUE)


ggplot(data = gc_df, aes(x = x, y = y)) + geom_raster(aes(fill = gc_dem)) +
  coord_quickmap() +
  theme_bw() + 
  scale_fill_gradientn(colors = c(
    "purple", "magenta", "red", "orange", "yellow", "white")
  )

```

How to select cell from a criteria
```{r}
gc_hab <- gc_crop

# Set any cells outside of [1000 - 15000] to NA
gc_hab[gc_hab > 1500 | gc_hab < 1000] <- NA

plot(gc_hab)
```


Now lets make this interactive with tmap
```{r}
tmap_mode("view")

tm_shape(gc_hab) + 
  tm_raster(legend.show = FALSE, palette = "plasma")
```

Kringing Rain in Kansas

Read in KS counties shapefile data
```{r}
ks_counties <- read_sf(here("data", "ks_counties", "ks_counties_shapefile.shp"))

# Use base plot
plot(ks_counties)

# Check CRS
st_crs(ks_counties)

# Set CRS
st_crs(ks_counties) <- 4326

crs(ks_counties)

plot(ks_counties)
```

Practice in ggplot
```{r}
ggplot(data = ks_counties) + geom_sf()
```

Read in Rainfall Data
```{r}
ks_rain <- read_csv(here("data", "ks_rain.csv")) %>%
  clean_names
```

Update ks_rain data to be recognized as spatil points
```{r}
ks_sf <- st_as_sf(ks_rain, coords = c("lon", "lat"), crs = 4326)
```

```{r}
ggplot() +
  geom_sf(data = ks_counties)+
  geom_sf(data = ks_sf, 
          aes(color = amt, size = amt),
          show.legend = FALSE)
```

Kriging to Predict Rainfall
```{r}
ks_sp <- as_Spatial(ks_sf)
class(ks_sp)
```

Make a spatial pixels grid to make predictions over
```{r}
bbox(ks_sp)
lat <- seq(37, 40, length.out = 200)
long <- seq(-94.63, -101.75, length.out = 200)

# Make into a spatial grid
grid <- expand.grid(lon = long, lat = lat)
grid_sf <- st_as_sf(grid, coords = c("lon", "lat"), crs = 4326)
grid_sp <- as_Spatial(grid_sf)

plot(grid_sp)
```

 
 
 
Make a Variogram
```{r}
ks_vgm <- variogram(amt ~ 1, data = ks_sp)
plot(ks_vgm)

#My estimates for variogram parameters
# nugget = 0.2
# sill = 0.8
# range = 200

ks_vgm_fit <- fit.variogram(ks_vgm, model = vgm(nugget = 0.2, psill = 0.8, range = 200, model = "Sph"))
plot(ks_vgm, ks_vgm_fit)
ks_vgm_fit
```

 
 Now, kriging! 
```{r}
ks_krige <- krige(amt ~ 1, ks_sp, grid_sp, model=ks_vgm_fit)

spplot(ks_krige, "var1.pred")

```

Let's get it back into a format we're used to (data frames & sf objects):
```{r}
# Make a data frame from kriged predictions:
ks_df <- data.frame(ks_krige@data["var1.pred"], ks_krige@data["var1.var"],
                    ks_krige@coords) %>% 
  rename(longitude = coords.x1,
         latitude = coords.x2)

# Convert to sf object: 
rain_sf <- st_as_sf(ks_df, coords = c("longitude","latitude"))
st_crs(rain_sf) <- 4326


ggplot(rain_sf) + geom_sf(aes(color = var1.pred))
```

```{r}

# Get Kansas outline to crop: 
ks <- read_sf(dsn = here("data","states"), 
              layer = "cb_2017_us_state_20m") %>% 
  dplyr::select(NAME) %>% 
  filter(NAME == "Kansas") %>% 
  st_transform(crs = 4326)

plot(ks)

```

```{r}
# Find the intersection of the two: 
rain_sf_ks <- st_intersection(rain_sf, ks)

ggplot(rain_sf_ks) +
  geom_sf(aes(color = var1.pred)) +
  scale_color_gradientn(colors = c("white","green","magenta","purple")) +
  theme_bw()

```

 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
